---
title: "p8105_hw3_kh3074"
output: github_document
date: 10-14-2021
author: KaiYu He(kh3074)
---
```{r}
library(tidyverse)
library(p8105.datasets)
```


# Problem 1

## 1,Load Data
```{r}
data("instacart")
summary(instacart)
```
### interpretation of the data  
- The ***instacart*** data has `r nrow(instacart)` rows of data with `r ncol(instacart)` columns of features.
- The is an online grocery shopping data set, ***order_id*** is the identifier for each order, ***prodect_id*** is the product identifier, ***reordered*** equals 1 if the product is ordered by the user in the past, ***order_dow*** is the day of the week on which the product was placed, ***days_since_prior_order*** is days since the last order, capped at 30, NA if order_number=1.All the orders are selected from eval_set == train.  
- Take the first line as an example, Bulgarian Yogurt was placed by user '112108' from 'dairy eggs' department. The user has already bought this item before.  

### 2
```{r}
instacart%>%
  count(aisle)%>% 
  arrange(desc(n))%>%
  head()%>%
  knitr::kable() #show the most 5 items ordered
```

- There are `r instacart%>%count(aisle)%>%nrow()` aisles in the data, accroding the data frame, ***fresh vegetable*** is the most ordered item.  


### 3

```{r}
instacart%>%
  count(aisle)%>%
  filter(n>10000)%>%
  mutate(aisle = fct_reorder(aisle,n))%>%
  ggplot(aes(x = n,y = aisle))+
  geom_point()

```

### 4  
```{r}
instacart%>%
  filter( aisle == 'baking ingredients'|aisle == 'dog food care'|aisle == 'packaged vegetables fruits')%>%
  group_by(aisle,product_name)%>%
  summarize(
    n_obs = n()
  )%>%
  top_n(3)%>%
  knitr::kable()
```

### 5
```{r}
instacart%>%
  filter(product_name == 'Pink Lady Apples'|product_name == 'Coffee Ice Cream')%>%
  group_by(order_dow,product_name)%>%
  summarize(
    mean_hour_of_day = mean(order_hour_of_day)
  )%>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour_of_day
  )%>%
  knitr::kable()
```

# Problem 2  

## 1,Load data  
```{r}
data("brfss_smart2010")

brfss_smart2010%>%
  janitor::clean_names()%>% #format the data to use appropriate variable names;
  filter(topic == 'Overall Health') #focus on the “Overall Health” topic
  
```

## 2,Data Cleaning  

```{r}
brfss_smart2010 = 
  brfss_smart2010%>%
  janitor::clean_names()%>% #format the data to use appropriate variable names;
  filter(topic == 'Overall Health')%>% #focus on the “Overall Health” topic
  filter(response == 'Excellent'|response == 'Very good'|response == 'Good'|response == 'Fair'|response == 'Poor')%>%
  #include only responses from “Excellent” to “Poor”
  mutate(rank = response)%>%
  mutate(rank = recode(rank,'Excellent' = 5,'Very good' = 4,'Good' = 3,'Fair' = 2, 'Poor' = 1))%>%
  mutate(response = as.factor(response))%>%
  mutate(response = fct_reorder(response,rank))%>% #organize responses as a factor taking levels ordered from “Poor” to “Excellent”
  select(-rank)
```

## 3  
```{r}
state_2002 = 
  brfss_smart2010%>%
  filter(year == 2002)%>%
  group_by(locationabbr)%>%
  rename('state_2002' = locationabbr)%>%
  summarize(locations_obversed = n()/5)%>% #Five levels for each location
  filter(locations_obversed >= 7)
state_2002%>%
  knitr::kable()
```

- According to the table, there are `r nrow(state_2002)` states were observed at more than 7 locations.
  
```{r}
state_2010 = 
  brfss_smart2010%>%
  filter(year == 2010)%>%
  group_by(locationabbr)%>%
  rename('state_2010' = locationabbr)%>%
  summarize(locations_obversed = n()/5)%>% #Five levels for each location
  filter(locations_obversed >= 7)
state_2010%>%
  knitr::kable()
```

- According to the table, there are `r nrow(state_2010)` states were observed at more than 7 locations.

## 4   
```{r}
brfss_smart2010%>%
  filter(response == 'Excellent')%>%
  group_by(year,locationabbr)%>%
  summarize(mean_excellent_value = mean(data_value))%>%
  ggplot(aes(x = year,y = mean_excellent_value,color = locationabbr))+
  geom_line(alpha = 0.5)
```

## 5    
```{r}
brfss_smart2010%>%
  filter(year == 2006|year == 2010)%>%
  filter(locationabbr == 'NY')%>%
  group_by(year,response)%>%
  ggplot(aes(y = data_value,x = locationdesc,color = response))+
  geom_point()+
  coord_flip()+
  facet_grid(~year)
```
# Problem 3  

## 1,Load and tidy data
```{r}
accel = read_csv("accel_data.csv")
accel_tidy = 
  accel%>%
  pivot_longer(
    activity.1:activity.1440,
    names_to = 'minute_of_day',
    values_to = 'number_of_activity' #merge the acctivity per minuter to a single column
  )%>%
  mutate(minute_of_day = str_replace(minute_of_day,'activity.',''))%>% 
  mutate(minute_of_day = as.numeric(minute_of_day))%>% # translate minute variable to numeric
  mutate(weekday_or_weekend = ifelse(day %in% c('Saturday','Sunday'),'weekend','weekday'))
accel_tidy
```
- There are total `r nrow(accel_tidy)` rows of obs in the tidy dataset, which contain number of activities per minute(1440 per day) in `r nrow(accel)` days.

## 2 Sum up
```{r}
accel_tidy%>%
  group_by(day_id)%>%
  summarize(activities_per_day = sum(number_of_activity))%>%
  knitr::kable()
accel_tidy%>%
  group_by(day_id)%>%
  summarize(total_number_activities = sum(number_of_activity))%>%
  ggplot(aes(x = day_id,y = total_number_activities))+
  geom_point()+
  geom_line()
```
- From the table and the plot, we can tell that the total activity over the day don't have a clear trend, it's more likely to be a white noise sequence.

## 3 plot
```{r}
# plot in minute
accel_tidy%>%
  mutate(day_id = as.character(day_id))%>%
  ggplot(aes(x = minute_of_day,y = number_of_activity,color = day))+
  geom_line(alpha = 0.5)

# plot in hour
accel_tidy%>%
  mutate(day_id = as.character(day_id))%>%
  mutate(hour_of_day = ceiling(minute_of_day/60))%>%
  group_by(day_id,hour_of_day)%>%
  summarize(
    numberof_activities_hour = sum(number_of_activity),
    day = day
  )%>%
  ggplot(aes(x = hour_of_day, numberof_activities_hour ,color = day))+
  geom_line(alpha = 0.5)

#plot divided by weekday and weekend
accel_tidy%>%
  mutate(day_id = as.character(day_id))%>%
  mutate(hour_of_day = ceiling(minute_of_day/60))%>%
  group_by(day_id,hour_of_day)%>%
  summarize(
    numberof_activities_hour = sum(number_of_activity),
    weekday_or_weekend = weekday_or_weekend
  )%>%
  ggplot(aes(x = hour_of_day, numberof_activities_hour ,color = weekday_or_weekend))+
  geom_line(alpha = 0.5)

```


- From the activities per day in minute and in hour, we can find out that, activities before 5 am and realy stable, within 1~4 pm  is relatively low.
- number of activities in 8~12 am is relatively high
- 20~22 pm is the peak of activities for every day.
- The curve in the plot fits people's daily routine very well.
- From the weekday vs weekend plot we can find out that during weekdays, 6~10 am and 8~10 pm are the peak of activities, during weekend the number of activities between 10~20 is relatively higher than weekdays.
